Impact offsetting

### 2022-03-28 paper.tex
Trajectory vs policy thoughts

		There is an obvious degree of freedom in the counterfactual question,
		even before we get to specifying what the generator should be doing.
		When should the counterfactual intervention happen? The answer to this
		question is the main difference between Krakovna's and Turner's approaches
		-- Krakovna answers ``at the beginning of time'', while Turner ``just a step ago''.
		Both these answers introduce some problems. I would argue we should be looking at
		a discounted combination of counterfactual interventions, e.g. rollout using the baseline
		policy starting at $4^n$ steps ago for $n \in \left\{ 0, \dots, k \right\}$.
		We should discount the penalties related to these counterfactuals by $2^k$.
		This avoids the tractability issues with Krakovna's approach, while avoiding the problem
		with Turner immediately accepting any loss in reversability after the agent is one step away
		form the decision that caused it.

		(I have a brilliant thought about this corresponding to the sunk cost fallacy, but
		not enough energy to write this out properly. Soemthingsomething
		only caring about the last decision makes you miss that your whole policy got you
		into deep trouble.)

		The generator itself could be trained on random policies with baseline rollouts
		computed starting from past states. Note that this requires the baseline to
		necessarily be a complete policy, being able to act in any state. This might
		be the inaction policy, but that would likely be unsatisfactory as a safe choice
		in general.
	\subsection{CVAE}
		A relatively natural approach to the above problem is to use a slightly modified CVAE.
		We create a rollout of a random policy $\pi$, of the form $\left( s_0, \dots s_T \right)$.
		We also create rollouts of the baseline policy $\pi_b$ of the forms
		\begin{equation*}
			\left( s^{\left( m \right)}_0 = s_m, \dots, s^{\left( m \right)}_{T-m} \right).
		\end{equation*}
		For every $t$ we use $s_t$ as an input to the CVAE, backpropagating a combination of
		$k$ losses
		\begin{equation*}
			L(s_t, s^{\left( t-1 \right)}_{1}, s^{\left( t-4 \right)}_4, \dots, s^{\left( t-4^k \right)}_{4^k}) = \sum_{l = 0}^k 2^{-l}L_{CVAE}\left( s_t, s^{\left( t-4^l \right)}_{4^l} \right)
		\end{equation*}
		Where $L_{CVAE}$ is the standard CVAE loss.

>>>>

		There is an obvious degree of freedom in the counterfactual question,
		even before we get to specifying what the generator should be doing.
		When should the counterfactual intervention happen? The answer to this
		question is the main difference between Krakovna's and Turner's approaches
		-- Krakovna answers ``at the beginning of time'', while Turner ``just a step ago''.
		Both these answers introduce some problems. I would argue we should be looking at
		a discounted combination of counterfactual interventions, i.e.
		\begin{equation}
			\baseline\left( s_T \right) = \frac{1-\gamma}{\gamma} \sum_{i=1}^T \gamma^{i} \baseline_{s_{T-i}}\left( s_T \right)
			\label{eq:discountedBaseline}
		\end{equation}
		where $\baseline_s$ corresponds to taking state $s$ as the point of counterfactual intervention.
		In this language Krakovna's baseline is equal to $\baseline_{s_0}$ and Turner's to $\baseline_{s_{T-1}}$.
		For small-scale experiments these baselines could just be results of parallel rollouts,
		but generators for them, or even the combined baseline, can in principle also be trained.
	\subsection{Advantages -- policy vs trajectory baseline}
	 This approach lessens the problem with Turner immediately accepting any loss in reversability
		after the agent is one step away from the decision that caused it. Such losses still
		become less relevant over time, but in a much more gradual way.

		The comparison with Krakovna is more involved. This approach definitely answers some problems
		related to what ``beginning of episode'' might mean, but in itself it's not a strong reason to
		favour it. For a better reason one has to notice that Krakovna \emph{de facto} uses a baseline
		\emph{trajectory} rather than a baseline \emph{policy}. To illustrate why this distinction
		matters let us discuss an environment:
		\begin{example}[Charging stations]
			Consider a gridworld in the shape of a 5 by 5 square. Points $\left( 0,0 \right)$,
			$\left( 0,2 \right)$, $\left( 4,0 \right)$ contain ``charging stations''. The agent is tasked
			with picking up pieces of trash randomly appearing in other places on the map.
		\end{example}
		This is only a very basic example to illustrate the problems. Also note that it is stochastic --
		this could be avoided, but the added complications would just muddy the waters.

		A natural safe baseline for this environment consists of moving to the closest charging station
		and staying there. However, if the agent originally starts in a specific point (say, $\left( 0,0 \right)$),
		an agent following Krakovna's approach will be incentivized to always go to that point.
		The baseline trajectory is always at this point, and the discounting in the future task measure
		means that the agent will be comparatively penalized for staying at any other point.

		The natural idea to fix it, randomizing the starting state, doesn't quite work. Since the starting states
		are not symmetric, a uniform distribution over the charging stations introduces at least an incentive
		to stay away from point $\left( 4,0 \right)$. A uniform distribution over all points is harder
		to analyse, but the way the resulting behaviour depends on the exact shape of the environment
		means it will be impossible to consistently recover the intended baseline policy.

		In contrast Turner's approach looks at the last state, thus picking behaviour that makes sense locally.
		Similarly the discounted combined baseline, although it has some lingering effects from states further
		back in time.

		This problem in general happens for any situation in which the dynamic system induced by the baseline
		policy has a set of attractors different from a single point. Then any single trajectory ends up
		in one of these attractors, and incentives in Krakovna's approach are skewed towards going to that
		attractor. Another relatively natural example of such situation is a baseline with a patrol route ––
		in Krakovna's approach the agent would be penalized for being out of sync with the patrol
		corresponding to the starting state.

		Krakovna mentioned that extending her methods to stochastic environments is one possible
		avenue of future research. She wasn't quite talking about this specific problem, but I suspect
		the underlying reason for problems in stochastic environments might be the same. In particular,
		the example she uses also talks about stochastically combining trajectories being unsatisfactory,
		and I believe the combined baseline approach should actually work in the sometimes-human-opens-window
		environment.

	\subsection{CVAE}
		The generator for the combined baseline could be trained on random policies with
		baseline policy rollouts	computed starting from past states. For tractability it might
		be better to sample fewer point baselines, e.g. only one every $4^k$ steps. This requires
		some rescaling of the discount factors which will be left as an exercise to the implementer.

		A relatively natural approach to this problem is to use a slightly modified CVAE.
		We create a rollout of a random policy $\pi$, of the form $\left( s_0, \dots s_T \right)$.
		We also create rollouts of the baseline policy $\pi_b$ of the forms
		\begin{equation*}
			\left( s^{\left( m \right)}_0 = s_m, \dots, s^{\left( m \right)}_{T-m} \right).
		\end{equation*}
		For every $t$ we use $s_t$ as an input to the CVAE, backpropagating a combination of
		$k$ losses
		\begin{equation*}
			L(s_t, s^{\left( t-1 \right)}_{1}, s^{\left( t-4 \right)}_4, \dots, s^{\left( t-4^k \right)}_{4^k}) = \sum_{l = 0}^k 2^{-l}L_{CVAE}\left( s_t, s^{\left( t-4^l \right)}_{4^l} \right)
		\end{equation*}
		Where $L_{CVAE}$ is the standard CVAE loss.
